Species motif participation provides unique information about species risk of extinction: supplemental information

### 2023-06-04 manuscript/supplement_motifs.tex
SI ready to submit

\title{Species motif participation provides unique information about species risk of extinction}

@@

            \begin{itemize}
                \item Describes GLMs relating persistence and other network properties (Eqn. S2-3) and GLMs relating motif participation, profiles and network properties (Eqn. S4-6)
            \end{itemize}

        \item Motif-persistence relationships vary with S,C

            \begin{itemize}
                \item Describes GLMs summarising how the relationship between motif participation and persistence varies over S and C, to complement the within-network results presented in the main text (Eqn. S7)
                \item Plots of these GLMs for each motif (Figs. S2-5)
            \end{itemize}

        \item Results: species motif participation 
            \begin{itemize}
                \item Supplemental results relating species motif participation to persistence and to other network properties (S, C, in-degree, STL)
                \item Table S1-2l; Fig. S6
            \end{itemize}

        \item Supplemental results: persistence and other network properties

            \begin{itemize}
                \item Supplemental results identifying relationships between persistence and network size, connectance, in-degree, and STL to provide context for motif relationships (Tables S3-4).
                \item Also relating average persistence to network motif profiles (Table S5)
            \end{itemize}

@@

            \begin{itemize}
                \item Relating motif profiles  to network size and connectance (Table S6)
            \end{itemize}

@@

        We simulated a suite of food webs based on the probabilistic niche model, which assigns predator-prey links based on the body-mass ratios between individuals of different species~\citep{Williams2000,Delmas2017}. 
        The meso-scale structure of niche-model networks closely mimics that of empirical food webs~\citep{Stouffer2007}. 
        To ensure that we captured a variety of realistic community sizes and structures, we generated networks ranging between 50 and 100 species (in steps of 10) with connectance values between 0.02 and 0.2 (in steps of 0.02). 
        The range of network sizes was chosen to reflect moderately well-sampled empirical webs while working within our computational limits, while the range of connectance values was chosen to cover that observed in most empirical food webs~\citep{Dunne2002e}.
        We generated a total of 100 networks with each combination of parameters, for a total of 6000 networks. 
        All networks were generated using the function ``nichemodel'' within the Julia language package \emph{BioEnergeticFoodWebs}~\citep{bioenergeticfw,Delmas2017}. 
        If a simulated network contained any disconnected species (species without predators or prey) or disconnected components (a group of species connected among themselves but not to the rest of the network), the network was rejected and a new network simulated.
        Finally, networks where the path lengths between each species and a basal resource could not be resolved (i.e., trophic levels were undefined) were rejected and new networks simulated.

@@

    \subsection{Filtering of biologically unrealistic networks}

        After simulating our initial niche-model networks, we filtered networks for biological plausibility.
        Specifically, we removed any network containing disconnected components (species or groups of species not connected to the rest of the network) or any network where any species had a shortest trophic level \textgreater6 as such high trophic levels are very uncommon in empirical food webs ~\citep{Riede2011}.
        Any removed network was replaced with a fresh simulation and re-checked until 100 realistic networks for each size-connectance combination had been obtained.

@@

\section{Motif-persistence relationships vary with S,C}

@@

    \begin{figure}[hb!]
        \centering
        \includegraphics[width=\textwidth]{figures/persistence_apparent_detailpers.eps}
        \caption{Slopes of the motif proportion-persistence relationships for the apparent competition motif were steepest in large and less-connected networks but the directions of slopes varied only with the strength of disturbance.}
        \label{apparent_plus_SC}
        \end{figure}

    \begin{figure}[hb!]
        \centering
        \includegraphics[width=\textwidth]{figures/persistence_direct_detailpers.eps}
        \caption{Slopes of the motif proportion-persistence relationships for the direct competition motif were steepest in large and less-connected networks but the directions of slopes varied only with the strength of disturbance.}
        \label{direct_plus_SC}
        \end{figure}

@@

    \begin{figure}[ht!]
        \centering
        \includegraphics[width=\textwidth]{figures/roles_vs_SC_all.eps}
        \caption{The proportion of each motif in consumer's participation vectors varied with connectance, network size, and their interaction (except for apparent competition).}
        \label{fig:motifs_vs_TL_and_deg}
    \end{figure}

@@

\section{Supplemental results: persistence and other network properties}

@@

    % Updated with network-level random effect
    \begin{table}[hb!]
        \caption{Coeffcients ($\beta$) for general linear mixed-effect models of consumers' probabilities of persistence against simple measures of species' role (STL or in-degree), level of disturbance to basal resources, and their interaction, as well as a random effect of global network properties (size and connectance). The model for STL was singular when including the random effect, so we removed this term and re-fit the model as a general linear model with fixed effects only. All models had binomial error distribution. Predictors were centered and scaled before model fitting. All $p$-values were \textless0.001. The level of disturbance is reflected by basal resources having a baseline probability of extinction ranging between $0.1 - 0.5$.  We also provide the marginal $R^2_M$ (fixed effects only) and conditional $R^2_C$ (fixed and random effects) for the in-degree model and the pseudo-$R^2$ for the STL model.}
        \label{tab:per_vs_TLdeg}
        \centering
        \begin{tabular}{l|c  c  c | c c  }
            Simple role measure & Role & Disturbance & Interaction & R$^2_M$ & R$^2_C$ \\
            \hline
            STL & \textbf{-0.0977} & \textbf{-0.600} & \textbf{0.0302} & \textbf{0.920} & NA \\
            in-degree & \textbf{-1.30$\times10^{-3}$} & \textbf{-0.139} & \textbf{-9.08$\times10^{-3}$} & 0.903 & 0.914 \\
        \end{tabular}
    \end{table}

@@

    % Updated to use mean instead of all species
    \begin{table}[ht!]
        \caption{These results correspond to equation 2, main text. Coefficients ($\beta$) for models relating persistence to the proportion of a network profile made up by a focal motif, the level of disturbance to basal resources (corresponding to a probability of extinction ranging between 0.1 and 0.5), and their interaction. Coefficients in \textbf{bold} are significant ($\alpha$\textless0.005). Predictors were centered and scaled before model fitting; coefficients refer to a unit increase in predictors. We also provide McFadden's pseudo-$R^2$ value for each regression. }
        \label{motif_profile_tab}
        \centering
        \footnotesize
        \begin{tabular}{c|c c c c c | c }
        Motif & Intercept & Proportion & Disturbance & Interaction &  pseudo-$R^2$ \\
            \hline
            Three-species chain & \textbf{0.326} &  0.0260 & \textbf{-0.595} & -2.53$\times10^{-3}$ & 0.958 \\
            Apparent competition & \textbf{0.326} & \textbf{0.0501} & \textbf{-0.596} & -2.00$\times10^{-3}$ & 0.962 \\
            Direct competition & \textbf{0.326} & 0.0151 & \textbf{-0.595} & -5.34$\times10^{-4}$ & 0.956 \\
            Omnivory & \textbf{0.326} & \textbf{-0.0626} & \textbf{-0.596} & 3.22$\times10^{-3}$ & 0.966 \\
        \end{tabular}
    \end{table}

>>>>

\title{Species motif participation provides unique information about species risk of extinction: supplemental information}

@@

            \begin{itemize}
                \item GLMs relating persistence and other network properties (Eqn. S2-S3)
                \item GLMs relating motif participation, profiles and network properties (Eqn. S4-S6)
            \end{itemize}

        \item Motif-persistence relationships vary with S, C

            \begin{itemize}
                \item GLMs summarising how the relationship between motif participation and persistence varies over S and C, to complement the within-network results presented in the main text (Eqn. S7)
                \item Plots of these GLMs for each motif (Figs. S2-S5)
            \end{itemize}

        \item Supplemental results: motif participation 
            \begin{itemize}
                \item Supplemental results relating species motif participation to persistence and to other network properties (S, C, in-degree, STL)
                \item Table S1-S2; Fig. S6
            \end{itemize}

        \item Supplemental results: persistence

            \begin{itemize}
                \item Supplemental results identifying relationships between persistence and network size, connectance, in-degree, and STL to provide context for motif relationships (Tables S3-S4).
                \item Supplemental results relating average persistence to network motif profiles (Table S5)
            \end{itemize}

@@

            \begin{itemize}
                \item Supplemental results relating motif profiles  to network size and connectance (Table S6; FIg. S7)
            \end{itemize}

@@

        We simulated a suite of food webs based on the probabilistic niche model, which assigns predator-prey links based on the body-mass ratios between individuals of different species~\citep{Williams2000,Delmas2017}. 
        The meso-scale structure of niche-model networks closely mimics that of empirical food webs~\citep{Stouffer2007}. 
        To ensure that we captured a variety of realistic community sizes and structures, we generated networks ranging between 50 and 100 species (in steps of 10) with connectance values between 0.02 and 0.2 (in steps of 0.02). 
        The range of network sizes was chosen to reflect moderately well-sampled empirical webs while working within our computational limits, while the range of connectance values was chosen to cover that observed in most empirical food webs~\citep{Dunne2002e}.
        We generated a total of 100 networks with each combination of parameters, for a total of 6000 networks. 
        All networks were generated using the function ``nichemodel'' within the Julia language package \emph{BioEnergeticFoodWebs}~\citep{bioenergeticfw,Delmas2017}. 
        If a simulated network contained any disconnected species (species without predators or prey) or disconnected components (a group of species connected among themselves but not to the rest of the network), the network was rejected and a new network simulated.
        Finally, networks where the path lengths between each species and a basal resource could not be resolved (i.e., trophic levels were undefined) were rejected, as was any network where any species had a shortest trophic level \textgreater6 as such high trophic levels are very uncommon in empirical food webs ~\citep{Riede2011}.
        Rejected networks were replaced with fresh simulations.

@@

@@

\section{Motif-persistence relationships vary with S, C}

@@

    \begin{figure}[hb!]
        \centering
        \includegraphics[width=\textwidth]{figures/persistence_apparent_detailpers.eps}
        \caption{Slopes of the motif proportion-persistence relationships for the apparent competition motif were steepest in large and less-connected networks or small and highly-connected networks but the directions of slopes varied only with the strength of disturbance.}
        \label{apparent_plus_SC}
        \end{figure}

    \begin{figure}[hb!]
        \centering
        \includegraphics[width=\textwidth]{figures/persistence_direct_detailpers.eps}
        \caption{Slopes of the motif proportion-persistence relationships for the direct competition motif were steepest in less-connected networks but the directions of slopes varied only with the strength of disturbance.}
        \label{direct_plus_SC}
        \end{figure}

@@

    \begin{figure}[ht!]
        \centering
        \includegraphics[width=\textwidth]{figures/roles_vs_SC_all.eps}
        \caption{The proportion of each motif in consumer's participation vectors varied with connectance, network size, and their interaction (except for apparent competition, where there was no significant interaction).}
        \label{fig:motifs_vs_TL_and_deg}
    \end{figure}

@@

\section{Supplemental results: persistence}

@@

    % Updated with network-level random effect
    \begin{table}[hb!]
        \caption{Coeffcients ($\beta$) for general linear mixed-effect models of consumers' probabilities of persistence against simple measures of species' role (STL or in-degree), level of disturbance to basal resources, and their interaction, as well as a random effect of global network properties (size and connectance). The model for STL was singular when including the random effect, so we removed this term and re-fit the model as a general linear model with fixed effects only. All models had binomial error distribution. All $p$-values were \textless0.001. The level of disturbance is reflected by basal resources having a baseline probability of extinction ranging between $0.1 - 0.5$.  We also provide the marginal $R^2_M$ (fixed effects only) and conditional $R^2_C$ (fixed and random effects) for the in-degree model and the pseudo-$R^2$ for the STL model.}
        \label{tab:per_vs_TLdeg}
        \centering
        \begin{tabular}{l|c  c  c | c c  }
            Simple role measure & Role & Disturbance & Interaction & R$^2_M$ & R$^2_C$ \\
            \hline
            STL & \textbf{-0.0977} & \textbf{-0.600} & \textbf{0.0302} & \textbf{0.920} & NA \\
            in-degree & \textbf{-1.30$\times10^{-3}$} & \textbf{-0.139} & \textbf{-9.08$\times10^{-3}$} & 0.903 & 0.914 \\
        \end{tabular}
    \end{table}

@@

    % Updated to use mean instead of all species
    \begin{table}[ht!]
        \caption{These results correspond to equation 2, main text. Coefficients ($\beta$) for models relating persistence to the proportion of a network profile made up by a focal motif, the level of disturbance to basal resources (corresponding to a probability of extinction ranging between 0.1 and 0.5), and their interaction. Coefficients in \textbf{bold} are significant ($\alpha$\textless0.005). We also provide McFadden's pseudo-$R^2$ value for each regression. }
        \label{motif_profile_tab}
        \centering
        \footnotesize
        \begin{tabular}{c|c c c c c | c }
        Motif & Intercept & Proportion & Disturbance & Interaction &  pseudo-$R^2$ \\
            \hline
            Three-species chain & \textbf{0.326} &  0.0260 & \textbf{-0.595} & -2.53$\times10^{-3}$ & 0.958 \\
            Apparent competition & \textbf{0.326} & \textbf{0.0501} & \textbf{-0.596} & -2.00$\times10^{-3}$ & 0.962 \\
            Direct competition & \textbf{0.326} & 0.0151 & \textbf{-0.595} & -5.34$\times10^{-4}$ & 0.956 \\
            Omnivory & \textbf{0.326} & \textbf{-0.0626} & \textbf{-0.596} & 3.22$\times10^{-3}$ & 0.966 \\
        \end{tabular}
    \end{table}
